\textsf{Are frontier models superhuman chemists?}

### 2024-03-15 src/tex/appendix.tex
add command for chembench

\subsection{Related work}
Existing benchmarks such as those from \citet{guo2023large}, \cite{sun2023scieval}, \citet{Schulze_Balhorn_2024}, \citet{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}). 
While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
ChemBench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols. 
Moreover, we believe that our robust framework will engage the community in open-source contributions.

@@

\subsection{Model performance}
We also evaluated the model performance on the entire chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.

@@

\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the chembench corpus.

>>>>

\subsection{Related work}
Existing benchmarks such as those from \citet{guo2023large}, \cite{sun2023scieval}, \citet{Schulze_Balhorn_2024}, \citet{Cai_2024} fail to comply with most of the requirements stipulated above (see \Cref{tab:existing benchmarks}). 
While these benchmarks could provide useful insights in the short-term, they cannot follow the rapid additions to the \gls{llm} space. 
\chembench aims to correct this through a set of developments: compatibility with BigBench, end-to-end automation, special focus on chemical safety, employment of diverse prompting strategies and specialized notation for molecules and mathematical symbols. 
Moreover, we believe that our robust framework, including the platform \url{chembench.org}, will engage the community in open-source contributions.

@@

\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.

@@

\Cref{fig:all_questions_models_completely_correct_radar} shows the performance of the models on the different topics of the \chembench corpus.

### 2024-03-15 src/tex/ms.tex
add command for chembench

    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
    Here, we introduce \enquote{chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.

@@

In this work, we report a novel benchmarking framework, which we call chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also an tool-augmented systems).

@@

Importantly, our large pool of questions encompasses a wide range of topics.
This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
The distribution of topics is also evident from \Cref{fig:pca} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
In this representation, semantically similar questions are close to each other, and we color the points based on a classification into XX topics. 
It is clear that a focus of chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.

@@

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}

@@

Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
For instance, Galactica\cite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
To support this, chembench is designed to encode the semantic meaning of various parts of the question or answer.  
For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.

Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), chembench is designed to operate on text completions.
This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\cite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).

@@

% GFK performance -> would the models obtain a chemical license

% Calculation vs. no calculation
Given that some of the questions in chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.

@@

@@

@@

\section*{Data and code availability}
The code and data for chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{showyourwork!} framework.\cite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}. 
To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

>>>>

    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, including GPT-4, against the expertise of human chemists.

@@

In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for the use in the chemical sciences.
Our benchmark consists of \variable{output/total_number_of_questions.txt}\unskip question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
Our corpus large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also an tool-augmented systems).

@@

Importantly, our large pool of questions encompasses a wide range of topics.
This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
The distribution of topics is also evident from \Cref{fig:pca} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
In this representation, semantically similar questions are close to each other, and we color the points based on a classification into XX topics. 
It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:pca} appear as large distinct clusters.

@@

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions in a balanced way (\variable{output/mcq_questions.txt} \gls{mcq} questions vs. \variable{output/non_mcq_questions.txt} open-ended questions).
As one might expect, most questions are difficult to read according to the Flesch–Kincaid readability test (\variable{output/flesch_kincaid_reading_ease.txt}).\cite{kincaid1975derivation}

@@

Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
For instance, Galactica\cite{taylor2022galactica} uses special tokenization or encoding procedures for molecules or equations.
To support this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
For instance, molecules represented in \gls{smiles} are encapsulated in \texttt{[START\_SMILES][\textbackslash END\_SMILES]}.

Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
This is also important with the growing number of tool-augmented systems that use external tools such as search \glspl{api} or code executors.\cite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\cite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application (e.g., in a copilot system for chemists).

@@

% GFK performance -> would the models obtain a chemical license 
% Interesting twist perhaps: the performance on the university textbooks or exams is better than on the ones we automatically constructed

% Calculation vs. no calculation
Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
For this, we manually annotated the questions that require calculations and performed the analysis on only those questions.

@@

% given that LLMs have parsed almost the entire internet, it is perhaps not surprising to seem them perform well on many of the task which often require 
% specific knowledge humans might not have at hand without consulting an additional knowledge bank 
% when it comes to reasoning tasks (NMR, point groups) the humans tend to beat LLMs

@@

Another aspect of requiring unsupervised question answering is that in real life humans have tools and are able to use them for answering any question of interest. 
In our current study, we prompted users to not use those tools.

@@

\section*{Data and code availability}
The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{showyourwork!} framework.\cite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}. 
To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
